Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-12-08 papers/integration/omm_integ.tex
More very small tweaks based on proofing

The \e{LINlinear} component, which is used in FEF, Thalamus, SC, SNr
and GPe populations, governs the activation $a$ with a first order
leaky integrator differential equation:
\begin{equation}
   \dot{a} = \frac{1}{\tau}(a_{in}-a)
\end{equation}
where $\tau$ is the time constant for the neural activation and
$a_{in}$ is the input to the neural element. $a_{in}$ is defined by an
activation input and a shunting inhibition input according to:
\begin{equation}
   a_{in} = A(1-s_a)+\alpha R_N
\end{equation}
Here, $A$ is the activation input and $s_a$ is the shunting inhibition
state variable whose value is related to the shunting input, $S$ by
\begin{equation}
   s_a = \begin{cases}
      S & S\leq 1 \\
      1 & S > 1
   \end{cases}
\end{equation}
$R_N$ is a random number drawn from a standard normal distribution
($\sigma$=1, $\mu$=0) and introduces noise to the activation of the
neural element, with the parameter $\alpha$ controlling the noise
amplitude.

@@

The mapping from the retinotopic co-ordinates in the brain to
rotational co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of striate cortex [visual stimulus to
electrophysiological response---\cite{daniel_representation_1961,
talbot_physiological_1941}] and by \cite{ottes_visuomotor_1986} for
superior colliculus data [electrophysiological SC stimulus to eye
movement response---\cite{robinson_eye_1972}].
% Ottes et al. fit to stimulus-to-movement data, and give a
% movement-to-collicular-coordinate mapping in the paper.
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar coordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar coordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian
co-ordinates of the neural map. We use $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ to give Euler rotations for the retinal target
stimulus. Note also that the form of
Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

@@

The resulting weight maps (where the oblique maps are
1/10$^{\mathrm{th}}$ of the vertical maps, as described earlier) are
shown in Fig.~\ref{fig:weightmaps}.
%
\rvw{First, recall that the $r$ axis of the neural surface corresponds
to the amplitude of a saccade and the $\phi$ axis indicates the polar
direction of the saccade, as described in Sect.~\ref{sec:mapping} and
Fig.~\ref{fig:mapping}. Fig.~\ref{fig:weightmaps}(a) shows the weight
map for the muscle which rotates the eye to the left. As we modelled a
left eye, this actuates the lateral rectus muscle. The exponential
rise of Eq.~\ref{eq:weightmaps} [for experimental evidence,
see Figs.~7 \& 8 of \cite{herrero_influence_1998}] is seen in the
$r$ direction; as $r$ increases, so the connection strength to the SBG
channel rises exponentially. The connection strength is greatest along
the centre line, for a value of $\phi$ which corresponds to a purely
leftward movement. Note that $\phi$ is presented in neural
coordinates, and not in degrees or radians; $\mathrm{1}\leq\phi\leq\mathrm{50}$
corresponds to a range of 0\dg~to 360\dg; $\phi=\mathrm{38.5}$ corresponds to
movements left. The connection strength drops away sinusoidally as
$\phi$ moves away from the centre line at $\phi=\mathrm{38.5}$. In regions of
the map for which there is no leftward movement, that is, in the half
of the map which corresponds to any movement with a rightward
component, the `left' weight map is 0. Fig.~\ref{fig:weightmaps}(d)
shows the weight map for rightward movements, actuating the medial
rectus muscle of the eye. The line of maximum connection strength is
along $\phi=\mathrm{13.5}$. The map is a mirror of
Fig.~\ref{fig:weightmaps}(a), reflected about the line $\phi\mathrm{26}$. The
`left' and `right' weight maps are orthogonal; the non-zero region of
the `left' map is zero in the `right' map and vice
versa. Fig~\ref{fig:weightmaps}(b) \& (d) show the weight maps for
downward and upward eye movements; the `down' map activates the SBG
channel for the inferior rectus muscle, the `up' map activates the
superior rectus. Note that `down' is not orthogonal either to `left'
or `right' because a saccade down and left is achieved by
simultaneously activating both the lateral and inferior rectus
muscles. However, the `up' map \e{is} orthogonal to the `down' map and
spans the edges of the surface where $\phi$ rolls over from 1 to
50. The line of maximum connection strength for the `up' map is along
$\phi=\mathrm{1}$; for `down' $\phi=\mathrm{26}$. Based on the training described in
Sect.~\ref{sec:methods:integrating}, the maps driving the superior
oblique (`Z+') and inferior oblique (`Z-') muscles were set to
1/10$^{\mathrm{th}}$ of the `down' and `up' maps.}
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

@@

The relationship between latency and the target luminance is given in
Fig.~\ref{fig:latvsall}(e). This shows latency for a 100~ms gap, step
and 100~ms overlap conditions for a given fixation luminance of 0.2,
and a horizontally located target at $\theta_{y}^{t}=$-10\dg. For the
gap condition, we see very short latencies for luminances of about
0.75 and above.  Finally, the activity driving these express saccades
is initiated by high firing rates in the superficial layer of SC
(SCs), which then drives activity in thalamus and through the basal
ganglia. A gradual transition from express saccades to reflexive
saccades is observed as the contribution of the SCs becomes weaker and
the drive from FEF into the thalamo-cortical loop becomes necessary to
elicit a saccade.
%
A similar gradual transition, albeit for higher latencies is seen for
the step condition. At higher target luminances, the SCs has a greater
effect on the activity in the thalamo-cortical loop. However, the
activity in STN caused by the fixation luminance increases the latency
at all luminance values compared with the gap condition.
%
The overlap condition leads to increased latencies for luminances
below 2.5, but meets the step condition above this value, at which the
0.2 fixation luminance appears to have a negligible effect on the
system.

>>>>

The \e{LINlinear} component, which is used in FEF, Thalamus, SC, SNr
and GPe populations, governs the activation $a$ with a first order
leaky integrator differential equation:
\begin{equation}
   \dot{a} = \frac{1}{\tau}(a_{in}-a)
\end{equation}
where $\tau$ is the time constant for the neural activation and
$a_{in}$ is the input to the neural element. $a_{in}$ is defined by an
activation input and a shunting inhibition input according to:
\begin{equation}
   a_{in} = A(1-s_a)+\alpha R_N
\end{equation}
Here, $A$ is the activation input and $s_a$ is the shunting inhibition
state variable whose value is related to the shunting input, $S$ by
\begin{equation}
   s_a = \begin{cases}
      S & S\leq 1 \\
      1 & S > 1
   \end{cases}
\end{equation}
$R_N$ is a random number drawn from a standard normal distribution
($\sigma=\mathrm{1}$, $\mu=\mathrm{0}$) and introduces noise to the activation of the
neural element, with the parameter $\alpha$ controlling the noise
amplitude.

@@

The mapping from the retinotopic co-ordinates in the brain to
rotational co-ordinates of the stimulus/response was written down by
\cite{schwartz_spatial_1977,eric_l._schwartz_computational_1980}
for measurements of striate cortex [visual stimulus to
electrophysiological response---\cite{daniel_representation_1961,
talbot_physiological_1941}] and by \cite{ottes_visuomotor_1986} for
superior colliculus data [electrophysiological SC stimulus to eye
movement response---\cite{robinson_eye_1972}].
% Ottes et al. fit to stimulus-to-movement data, and give a
% movement-to-collicular-coordinate mapping in the paper.
We used the following statement of this mapping to introduce stimuli
into the `World' input population of the brain model:
\begin{equation}\label{eq:thetas_to_phi}
   \phi = \frac{W_{nfs}} {2 \pi}\arctan\left(\frac{\theta_{y}^{t}}{\theta_{x}^{t}}\right)
\end{equation}
\begin{equation}\label{eq:thetas_to_r}
   r = M_fE_2\,\ln\left(\frac{1}{E_2}\sqrt{{\theta_{x}^{t}}^2 + {\theta_{y}^{t}}^2}+1\right)
\end{equation}
Note that we use $r$ and $\phi$ as the co-ordinates on the `collicular
surface'. Schwartz uses $r$ and $\phi$ as the polar co-ordinates of the
retinal stimulus; Ottes et al.~use $r$ and $\phi$ as polar co-ordinates
for the eye movement response; both use $u$ and $v$ as the Cartesian
co-ordinates of the neural map. We use $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ to give Euler rotations for the retinal target
stimulus. Note also that the form of
Eqns.~\ref{eq:thetas_to_phi} \& \ref{eq:thetas_to_r} is slightly
different from that given in
\cite{ottes_visuomotor_1986} because our $\theta_{x}^{t}$ and
$\theta_{y}^{t}$ are not the polar co-ordinates used in that work.

@@

The resulting weight maps (where the oblique maps are
1/10$^{\mathrm{th}}$ of the vertical maps, as described earlier) are
shown in Fig.~\ref{fig:weightmaps}.
%
\rvw{First, recall that the $r$ axis of the neural surface corresponds
to the amplitude of a saccade and the $\phi$ axis indicates the polar
direction of the saccade, as described in Sect.~\ref{sec:mapping} and
Fig.~\ref{fig:mapping}. Fig.~\ref{fig:weightmaps}(a) shows the weight
map for the muscle which rotates the eye to the left. As we modelled a
left eye, this actuates the lateral rectus muscle. The exponential
rise of Eq.~\ref{eq:weightmaps} [for experimental evidence,
see Figs.~7 \& 8 of \cite{herrero_influence_1998}] is seen in the
$r$ direction; as $r$ increases, so the connection strength to the SBG
channel rises exponentially. The connection strength is greatest along
the centre line, for a value of $\phi$ which corresponds to a purely
leftward movement. Note that $\phi$ is presented in neural
co-ordinates, and not in degrees or radians; $\mathrm{1}\leq\phi\leq\mathrm{50}$
corresponds to a range of 0\dg~to 360\dg; $\phi=\mathrm{38.5}$ corresponds to
movements left. The connection strength drops away sinusoidally as
$\phi$ moves away from the centre line at $\phi=\mathrm{38.5}$. In regions of
the map for which there is no leftward movement, that is, in the half
of the map which corresponds to any movement with a rightward
component, the `left' weight map is 0. Fig.~\ref{fig:weightmaps}(d)
shows the weight map for rightward movements, actuating the medial
rectus muscle of the eye. The line of maximum connection strength is
along $\phi=\mathrm{13.5}$. The map is a mirror of
Fig.~\ref{fig:weightmaps}(a), reflected about the line $\phi=\mathrm{26}$. The
`left' and `right' weight maps are orthogonal; the non-zero region of
the `left' map is zero in the `right' map and vice
versa. Fig~\ref{fig:weightmaps}(b) \& (d) show the weight maps for
downward and upward eye movements; the `down' map activates the SBG
channel for the inferior rectus muscle, the `up' map activates the
superior rectus. Note that `down' is not orthogonal either to `left'
or `right' because a saccade down and left is achieved by
simultaneously activating both the lateral and inferior rectus
muscles. However, the `up' map \e{is} orthogonal to the `down' map and
spans the edges of the surface where $\phi$ rolls over from 1 to
50. The line of maximum connection strength for the `up' map is along
$\phi=\mathrm{1}$; for `down' $\phi=\mathrm{26}$. Based on the training described in
Sect.~\ref{sec:methods:integrating}, the maps driving the superior
oblique (`Z+') and inferior oblique (`Z-') muscles were set to
1/10$^{\mathrm{th}}$ of the `down' and `up' maps.}
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

@@

The relationship between latency and the target luminance is given in
Fig.~\ref{fig:latvsall}(e). This shows latency for a 100~ms gap, step
and 100~ms overlap conditions for a given fixation luminance of 0.2,
and a horizontally located target at $\theta_{y}^{t}=\mathrm{-10}$\dg. For the
gap condition, we see very short latencies for luminances of about
0.75 and above.  Finally, the activity driving these express saccades
is initiated by high firing rates in the superficial layer of SC
(SCs), which then drives activity in thalamus and through the basal
ganglia. A gradual transition from express saccades to reflexive
saccades is observed as the contribution of the SCs becomes weaker and
the drive from FEF into the thalamo-cortical loop becomes necessary to
elicit a saccade.
%
A similar gradual transition, albeit for higher latencies is seen for
the step condition. At higher target luminances, the SCs has a greater
effect on the activity in the thalamo-cortical loop. However, the
activity in STN caused by the fixation luminance increases the latency
at all luminance values compared with the gap condition.
%
The overlap condition leads to increased latencies for luminances
below 2.5, but meets the step condition above this value, at which the
0.2 fixation luminance appears to have a negligible effect on the
system.
